### 2016-02-24 mcmc.tex
notes to self
problems, a comment, and I like my expectations specific

Given this pdf $p(\pars)$, we can define \emph{expectation values} $E[\pars]$
for $\pars$ or for any quantity that can be expressed as a function $q(\pars)$
of $\pars$:
\begin{eqnarray}
E[\pars] &\equiv& \int \pars\,p(\pars)\,\dd \pars
\\
E[q] &\equiv& \int q(\pars)\,p(\pars)\,\dd \pars
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $\pars$ (all the parts of $\pars$ space in which $p(\pars)$ is finite) and the
integrals are multi-dimensional if $\pars$ is multi-dimensional.
These expectation values are the mean values of $\pars$ and $q(\pars)$ under
the pdf.  A good sampling---and really this is the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{\pars_k}_{k=1}^K$ the integrals get replaced
with sums over samples $\pars_k$:
\begin{eqnarray}
E[\pars] &\approx& \frac{1}{K}\,\sum_{k=1}^K \pars_k
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

@@

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E[q] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E[q] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below in \sectionname~\ref{sec:MH}) from
evaluations of $f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

\paragraph{visualization:}
HOGG:...Make triangle plots.  Great value!...  Show example here... Mentioned above; drop here or reiterate?

>>>>

Given this pdf $p(\pars)$, we can define \emph{expectation values}
$E_{p(\pars)}[\pars]$
for $\pars$ or for any quantity that can be expressed as a function $q(\pars)$
of $\pars$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\equiv& \int \pars\,p(\pars)\,\dd \pars
\\
E_{p(\pars)}[q(\pars)] &\equiv& \int q(\pars)\,p(\pars)\,\dd \pars
\quad ,
\end{eqnarray}
where again the integrals are implicitly definite integrals over the
entire domain
of $\pars$ (all the parts of $\pars$ space in which $p(\pars)$ is finite) and the
integrals are multi-dimensional if $\pars$ is multi-dimensional.
These expectation values are the mean values of $\pars$ and $q(\pars)$ under
the pdf.  A good sampling---and really this is the definition of a good
sampling---makes the sampling approximation to these integrals
accurate.
With a good sampling $\setof{\pars_k}_{k=1}^K$ the integrals get replaced
with sums over samples $\pars_k$:
\begin{eqnarray}
E_{p(\pars)}[\pars] &\approx& \frac{1}{K}\,\sum_{k=1}^K \pars_k
\\
E_{p(\pars)}[q(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\quad .
\end{eqnarray}
That is, a sampling is good when any expectation value of interest is
accurately computed via the sampling approximation.
The word ``accurately'' here translates into some kinds of theorems
about limiting behavior; the general idea is that the sampling
approximation becomes exact as $K$ goes to infinity.
The size $K$ of the sampling you need \emph{in practice} will depend
on the expectations you want to compute, and the accuracies you need.

@@

DFM says: I don't agree with this section. Using MCMC you \emph{are} sampling
from $p(\pars)$\ldots you \emph{can't} sample from $f(\pars)$.

When the sampling $\setof{\pars_k}_{k=1}^K$ is of one of these badly
normalized functions $f(\pars)$---as it usually will be---the
sampling-approximation expectation values are the expectation values
under the properly-normalized corresponding pdf, even though you might
\emph{never learn that normalization}.
When we run \MCMC\ sampling on $f(\pars)$, a function that differs from a
pdf $p(\pars)$ by some unknown normalization constant, then the sampling
permits computation of the following kinds of quantities:
\begin{eqnarray}
E_{p(\pars)}[q(\pars)] &\equiv& \frac{\int q(\pars)\,f(\pars)\,\dd \pars}{\int f(\pars)\,\dd \pars}
\\
E_{p(\pars)}[q(\pars)] &\approx& \frac{1}{K}\,\sum_{k=1}^K q(\pars_k)
\label{eq:expectationvalue}
\quad .
\end{eqnarray}
That is, the sampling can be constructed (as we will show below in \sectionname~\ref{sec:MH}) from
evaluations of $f(\pars)$ directly, and it permits you to compute expectation values
without ever requiring you to integrate either the numerator integral
or the denominator integral, both of which are generally
intractable.\note{As we will
  see, the intractability comes from various directions, but one is
  that the dimension of $\pars$ gets large in most realistic situations.
  Another is that the support of $p(\pars)$ tends to be, in normal
  inference situations, \emph{far smaller} than the full domain of
  $\pars$.
  That is, the pdf is at or very close to zero over all but some tiny
  and very hard-to-find part of the space.}

@@

\paragraph{visualization:}
HOGG: What plots do you make to diagnose problems?

HOGG: Plot parameters vs ``time'' in the chain.

HOGG: Plot a cut through the density function and the samples near that cut.

HOGG:...Make triangle plots.  Great value!...  Show example here... Mentioned above; drop here or reiterate?
